---
title: "Homework3"
author: "Ruoyu_Jia & Xi_Chen"
date: "2/24/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##1
$$
\phi(y_i-x_i^{\top}\beta_j;0,\delta^2)=\frac{1}{\sqrt{2\pi}~\delta}e^{-\frac{(y_i-x_i^{\top}\beta_j)^2}{2\delta^2}}
$$
$$
\sum_{i=1}^n\sum_{j=1}^mp_{ij}^{k+1}\{\log\pi_i+\log\phi(y_i-x_i^{\top}\beta_i;0,\delta^2)\}=\sum_{i=1}^n\sum_{j=1}^mp_{ij}^{k+1}\{\log(\frac{1}{\sqrt{2\pi}~\delta}e^{-\frac{(y_i-x_i^{\top}\beta_j)^2}{2\delta^2}})\}  
$$
$$=\sum_{i=1}^n\sum_{j=1}^mp_{ij}^{k+1}\log(\frac{1}{\sqrt{2\pi}}\pi_j)-\frac{1}{2}\sum_{i=1}^n\sum_{j=1}^mp_{ij}^{k+1}\log(\delta^2)-\frac{1}{2}\sum_{i=1}^n\sum_{j=1}^mp_{ij}^{k+1}\frac{(y_i-x_i^{\top}\beta_j)^2}{\delta^2}=I_1-\frac{1}{2}I_2-\frac{1}{2}I_3
$$
Firstly, only $I_3$ contains $\beta_j~, j = 1,...m$.~ We only need $\beta_j$ to minimize each $I_{ij}$
$$
I_{ij}=\sum_{i=1}^np_{ij}^{k+1}~\frac{(y_i-x_i^{\top}\beta_i)^2}{\delta^2}
$$
$$
\frac{\partial I_{ij}}{\partial \beta_j}=-\frac{2\sum_{i=1}^np_{ij}^{k+1}x_i(y_i-x_i^\top\beta_j)}{\delta^2}=0
$$
$$
\sum_{i=1}^np_{ij}^{k+1}x_ix_i^{\top}\beta_j=\sum_{i=1}^np_{ij}^{k+1}x_iyi
$$
$$
\beta_j=(\sum_{i=1}^nx_ix_i^{\top}p_{ij}^{k+1})^{-1}(\sum_{i=1}^nx_ip_{ij}^{k+1}y_i)
$$


Next, only $I_2$ and $I_3$ contains $\delta^2$, we need $\delta^2$ to minimize $(I_2+I_3)$
$$
\frac{\partial (I_2+I_3)}{\partial( \delta^2)}=\frac{\sum_{i=1}^n\sum_{j=1}^mp_{ij}^{k+1}}{\delta^2}-\frac{\sum_{i=1}^n\sum_{j=1}^m(y_i-x_i^{\top}\beta_j)^2}{\delta^4}=0
$$
$$
\delta^2=\frac{\sum_{i=1}^n\sum_{j=1}^m(y_i-x_i^{\top}\beta_j)^2}{\sum_{i=1}^n\sum_{j=1}^mp_{ij}^{k+1}}=\frac{\sum_{i=1}^n\sum_{j=1}^m(y_i-x_i^{\top}\beta_j)^2}{n}
$$
Lastly, only $I_1$ contains $\pi_j$
$$
I_1=-\frac{1}{2}\log(2\pi)~\sum_{i=1}^n\sum_{j=1}^mp_{ij}^{k+1}
$$
So we need to minimize $\sum_{i=1}^n\sum_{j=1}^mp_{ij}^{k+1}$. Note that $\pi_1+\pi_2+...+\pi_m=1$,
$$
L=\sum_{i=1}^n\sum_{j=1}^mp_{ij}^{k+1}-\lambda(\sum_{j=1}^m\pi_j-1)
$$
$\lambda$ is a Lagrange multipter,
$$
\pi_j=\frac{\sum_{i=1}^np_{ij}^{k+1}}{\sum_{i=1}^n\sum_{j=1}^mp_{ij}^{k+1}}=\frac{\sum_{i=1}^np_{ij}^{k+1}}{n}
$$

##2(a)
$$
\int_0^\infty2X^{\theta-1}e^{-x}~dx=2\Gamma(\theta)\int_0^\infty\frac{X^{\theta-1}e^{-x}}{\Gamma(\theta)}~dx=2\Gamma(\theta)Gamma(\theta,1)
$$
$$
\int_0^\infty X^{\theta-1/2}~e^{-x}~dx=\int_0^\infty X^{(\theta+1/2)^{-1}}~e^{-x}~dx=\Gamma(\theta+1/2)\int_0^\infty\frac{X^{\theta-1}e^{-x}}{\Gamma(\theta+1/2)}~dx=\Gamma(\theta+1/2)Gamma(\theta+1/2,1)
$$
$$
C\int_0^\infty (2X^{\theta-1}~X^{\theta-1/2})~e^{-x}~dx=C(2\Gamma(\theta)Gamma(\theta,1)+\Gamma(\theta+1/2)Gamma(\theta+1/2,1))=1
$$
$$
C=\frac{1}{2\Gamma(\theta)Gamma(\theta,1)+\Gamma(\theta+1/2)Gamma(\theta+1/2,1)}
$$
$$
g(x)=\frac{(2X^{\theta-1}~X^{\theta-1/2})e^{-x}}{2\Gamma(\theta)Gamma(\theta,1)+\Gamma(\theta+1/2)Gamma(\theta+1/2,1)}
$$
$$
=2C\Gamma(\theta)\frac{X^{\theta-1}e^{-x}}{\Gamma(\theta)}+C\Gamma(\theta+1/2)\frac{X^{\theta-1/2}e^{-x}}{\Gamma(\theta+1/2)}
$$
$$
=2C\Gamma(\theta)g_1(x)+C\Gamma(\theta+1/2)g_2(x)
$$
Therefore, g is a mixture of Gamma distributions, which are $Gamma(\theta,1)$ and $Gamma(\theta+1/2,1)$. the weights of them are $\frac{2\Gamma(\theta)}{2\Gamma(\theta)+\Gamma(\theta+1/2)}$ and $\frac{2\Gamma(\theta+1/2)}{2\Gamma(\theta)+\Gamma(\theta+1/2)}$

##2(b)

### step 1

sample k from {1,2} 

the prob of k=1 is $2*\Gamma(\theta)*C$

the prob of k=2 is $\Gamma(\theta+1/2)*C$


### step 2

if k=1, get 1 sample from $Gamma(1,\theta)$ 

if k=2, get 1 sample from $Gamma(1,\theta+1/2)$


### step 3

repeat step1 and step2 10000 times 
\newpage

```{r ,echo=FALSE}

g.sample <- function(theta,n){
  sample = rep(0, n)
  for (i in 1:n){
    a = sample(1:2,1,prob = c(2*gamma(theta),gamma(theta+1/2)))
    if (a==1){
      x = rgamma(1,theta)
      sample[i] = x
    } else {
      x = rgamma(1,theta+1/2)
      sample[i] = x
    }
  }
  return(sample)
}

sample = g.sample(5,10000)
theta = 5 

#plot
intvl=c(.1, 15)
n.p=10001
dg.obj = density(sample, from=intvl[1], to=intvl[2], n=n.p)
x = dg.obj$x
dg = dg.obj$y
g.val = 2*x^(theta-1)*exp(-x)+x^(theta-1/2)*exp(-x)
g <- function(x) {2*x^(theta-1)*exp(-x)+x^(theta-1/2)*exp(-x)}
c <- integrate(g, lower = 0, upper = Inf)$value
g.val = g.val/c
title1 <- paste("theta=",theta)
plot(ts(cbind(dg,g.val),start=intvl[1],
     deltat=diff(intvl)/(n.p-1)), main = c(title1),
     plot.type="single", col=c("blue", "red"),
     ylab="density", xlab="x")
legend("topright",legend=c('sample','g'),col=c('blue','red'),
       lty=1)

```

## 2(c)
$f(x)=\alpha g(x)$, $f(x)/g(x)=\alpha=\frac{\sqrt{4+x}~x^{\theta-1}e^{-x}}{(2x^{\theta-1}+x^{\theta-1/2})e^{-x}}$ = $\frac{\sqrt{4+x}}{2+\sqrt x}$ = $F(x)$

$$
Let~~F'(x)=0
$$
Then we have $x=4$. Therefore, $\alpha=\sqrt2/2$ and accepting probability is $\frac{\sqrt{4+x}}{\sqrt 2/2(2+\sqrt x)}$

### step 1
get 1 sample from function g 

### step 2
u ~ unif(0,1) 

if u < $f(x)/\alpha g(x)$ 

accept it, go to step3 

otherwise, reject it, return step1 

### step 3
<br>
repeat step1 and step2 10000 times 
<br>

```{r ,echo=FALSE}
f.sample <- function(theta, n=10000) {
  sample = rep(0, n)
  for (i in 1:n) {
    while (1) {
      x = g.sample(theta,1)
      if ((sqrt(2)/2)*(2+sqrt(x))*runif(1) < sqrt(4+x)) {
        sample[i] = x
        break
      }
    }
  }
  
  #plot
  intvl=c(.1, 15)
  n.p=10001
  df.obj = density(sample, from=intvl[1], to=intvl[2], n=n.p)
  x = df.obj$x
  df = df.obj$y
  f.val = sqrt(4+x)*x^(theta-1)*exp(-1*x)
  f <- function(x) {sqrt(4+x)*x^(theta-1)*exp(-1*x)}
  c <- integrate(f, lower = 0, upper = Inf)$value
  f.val = f.val/c
  title1 <- paste("theta=",theta)
  plot(ts(cbind(df,f.val),start=intvl[1],
       deltat=diff(intvl)/(n.p-1)),main = c(title1),
       plot.type="single", col=c("blue", "red"),
       ylab="density", xlab="x")
  legend("topright",legend=c('sample','f'),col=c('blue','red'),
         lty=1)

  return(sample)
}

sample = f.sample(5)

```


##3(a)
Let $g(x)=(1-\lambda)\theta x^{\theta-1}+\lambda\beta(1-x)^{\beta-1}=(1-\lambda)g_\theta(x)+\lambda g_\beta(x), ~0<\lambda<1$. Here $(1-\lambda)~and~\lambda$ can be considered as the weight of $g_\theta~and~g_\beta$.
$$
\alpha=\sup(\frac{\frac{x^{\theta-1}}{1+x^2}+\sqrt{2+x^2}(1-x)^{\beta-1}}{(1-\lambda)\theta x^{\theta-1}+\lambda \beta(1-x)^{\beta-1}})
$$
As the expression of $\alpha$ is complicated, we can use scaling method to estimate the value of $\alpha$, which is
$$
\alpha\leq \sup(\frac{\frac{1}{1+x^2}+\sqrt{2+x^2}(1-x)^{\beta-1}x^{1-\theta}}{\lambda \beta(1-x)^{\beta-1}x^{1-\theta}})
$$
$$
=\sup(\frac{\frac{1}{1+x^2}}{\lambda \beta(1-x)^{\beta-1}x^{1-\theta}}+\frac{\sqrt{2+x^2}}{\lambda \beta})
$$
$$
=\frac{1}{\lambda\beta}\sup(\frac{(1-x)^{1-\beta}x^{\theta-1}}{1+x^2}+\sqrt{2+x^2})
$$
$$
\leq \frac{1}{\lambda\beta}\sup((1-x)^{1-\beta}x^{\theta-1}+\sqrt{2+x^2})
$$
$$
=\frac{\sqrt3}{\lambda\beta}
$$
Therefore, the accepting probability is $\frac{\frac{x^{\theta-1}}{1+x^2}+\sqrt{2+x^2}(1-x)^{\beta-1}}{\frac{\sqrt3}{\lambda\beta}((1-\lambda)\theta x^{\theta-1}+\lambda \beta(1-x)^{\beta-1})}$

### step 1
get 1 sample from function g 

### step 2

u ~ unif(0,1) 

if u < $f(x)/\alpha g(x)$ 

accept it, go to step3 

otherwise, reject it, return step1 

### step 3
repeat step1 and step2 10000 times 


```{r ,echo=FALSE}
f.sample <- function(theta,b, n=10000) {
  sample = rep(0, n)
  lambda = 0.6
  for (i in 1:n) {
    while(1){
      if (runif(1)>lambda){
        x = runif(1)^(1/theta)
      } else {
        x = 1-runif(1)^(1/b)
      }
      if (runif(1)*sqrt(3)/(lambda*b)*((1-lambda)*theta*x^(theta-1           )+lambda*b*(1-x)^(b-1))<x^(theta-1)/(1+x^2)+sqrt(2+x^2)*           (1-x)^(b-1)){
        sample[i] = x
        break
      }
    }
  }
  
  #plot
  intvl=c(0.05, 0.95)
  n.p=501
  df.obj = density(sample, from=intvl[1], to=intvl[2], n=n.p)
  x = df.obj$x
  df = df.obj$y
  f.val = x^(theta-1)/(1+x^2)+sqrt(2+x^2)*(1-x)^(b-1)
  f <- function(x) {x^(theta-1)/(1+x^2)+sqrt(2+x^2)*(1-x)^(b-1)}
  c <- integrate(f, lower = 0, upper = 1)$value
  f.val = f.val*mean(df)/c
  title1 <- paste("theta=",theta)
  title2 <- paste("beta=",b)
  plot(ts(cbind(df, f.val), start=intvl[1],                              deltat=diff(intvl)/(n.p-1)), main = c(title1, title2),
       plot.type="single", col=c("blue", "red"),
       ylab="density", xlab="x")
  legend("topright",legend=c('sample','f'),col=c('blue','red'),
         lty=1)
  
  return(sample)
}

sample1 = f.sample(5,5)
sample2 = f.sample(5,10)
```

##3(b)
$$
f(x)\propto  \frac{x^{\theta-1}}{1+x^2}+\sqrt{2+x^2}(1-x)^{\beta-1}=q_1(x)+q_2(x), 0<x<1
$$

$$
q_1(x)=\frac{x^{\theta-1}}{1+x^2}\leq\alpha_1g_1(x)=\alpha\frac{x^{\theta-1}}{B(\theta,1)}
$$
$$
\alpha=\sup(\frac{B(\theta,1)}{1+x^2})=B(\theta,1)
$$
So the accepting probability is $\frac{1}{1+x^2}$
$$
q_2(x)=\sqrt{2+x^2}~(1-x)^{\beta-1}\leq \alpha_2g_2(x)=\alpha_2\frac{(1-x)^{\beta-1}}{B(1,\beta)}
$$
$$
\alpha_2=\sup(B(1,\beta)\sqrt{2+x^2})=\sqrt3 B(1,\beta)
$$
So the accepting probability is $\frac{\sqrt{2+x^2}}{\sqrt3}$

### step 1    
sample k from {1,2} with probabilities $p_k \propto \alpha_k$ 

### step 2    
get 1 sample from $g_k(x)$    
<br>
u ~ unif(0,1) 
<br>
if u < $f_k(x)/\alpha g_k(x)$    
<br>
accept it    
<br>
otherwise, reject it, repeat step2 until 1 sample is accepted 

### step 3    
repeat step1 and step2 10000 times    


```{r ,echo=FALSE}
f.separate <- function(theta,b, n=10000) {
  sample = rep(0, n)
  for (i in 1:n) {
    a = sample(1:2,1,prob = c(beta(theta,1),sqrt(3)*beta(1,b)))
    if (a==1){
      while (1) {
        x = rbeta(1,theta,1)
        if ((1+x^2)*runif(1)<1) {
          sample[i] = x
          break
        } 
      }
    } else {
      while (1) {
        x = rbeta(1,1,b)
        if (sqrt(3)*runif(1)<sqrt(2+x^2)) {
          sample[i] = x
          break
        } 
      }
    }
  }
  
  #plot
  intvl=c(0.05, 0.95)
  n.p=10001
  df.obj = density(sample, from=intvl[1], to=intvl[2], n=n.p)
  x = df.obj$x
  df = df.obj$y
  f.val = x^(theta-1)/(1+x^2)+sqrt(2+x^2)*(1-x)^(b-1)
  f <- function(x) {x^(theta-1)/(1+x^2)+sqrt(2+x^2)*(1-x)^(b-1)}
  c <- integrate(f, lower = 0, upper = 1)$value
  f.val = f.val/c
  title1 <- paste("theta=",theta)
  title2 <- paste("beta=",b)
  plot(ts(cbind(df, f.val), start=intvl[1],                              deltat=diff(intvl)/(n.p-1)), main = c(title1, title2),
       plot.type="single", col=c("blue", "red"),
       ylab="density", xlab="x")
  legend("topright",legend=c('sample','f'),col=c('blue','red'),
         lty=1)
  
  return(sample)
}

sample3 = f.separate(5,5)
sample4 = f.separate(5,10)

```

